---
title: "scRNA_seq_root"
author: "Chenxin Li"
date: "5/17/2022"
output:  
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages 
```{r}
library(tidyverse)
library(Seurat) 
library(readxl)
library(RColorBrewer)
library(viridis)
library(svglite) 
library(patchwork)
library(cowplot)
```

# Load Data
```{r}
AW_raw <- Read10X(data.dir = "../Data/single_cell_matrix/AW/")
AX_raw <- Read10X(data.dir = "../Data/single_cell_matrix/AX")
```

# Set up Seurat object 
```{r}
AW <- CreateSeuratObject(counts = AW_raw, min.cells = 3, min.features = 200, project = "AW")
AX <- CreateSeuratObject(counts = AX_raw, min.cells = 3, min.features = 200, project = "AX")
```

# QC and cell selection
```{r}
VlnPlot(AW, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
VlnPlot(AX, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
```
```{r}
AW <- subset(AW, subset = nFeature_RNA > 200 & nFeature_RNA < 3000)
AX <- subset(AX, subset = nFeature_RNA > 200 & nFeature_RNA < 3000)
```

## Median number genes
```{r}
summary(AW$nFeature_RNA)
summary(AX$nFeature_RNA)
```
```{r}
summary(AW$nCount_RNA)
summary(AX$nCount_RNA)
```


# Normalize
```{r}
AW <- NormalizeData(AW, normalization.method = "LogNormalize",
                          scale.factor = 10000)

AX <- NormalizeData(AX, normalization.method = "LogNormalize",
                          scale.factor = 10000)
```

# Select feature
```{r}
AW <- FindVariableFeatures(AW, selection.method = "vst",
                                nfeatures = 3000)

AX <- FindVariableFeatures(AX, selection.method = "vst",
                                nfeatures = 3000)
```

```{r}
VariableFeaturePlot(AW)
VariableFeaturePlot(AX)
```
# Integration
```{r}
Root.list <- list(AW, AX)

R_features <- SelectIntegrationFeatures(
  object.list = Root.list, nfeatures = 3000
)

R_anchors <- FindIntegrationAnchors(object.list = Root.list,
                                    anchor.features = R_features)

Root_Integrated <- IntegrateData(anchorset = R_anchors) 
```

# Scale Data 
```{r}
DefaultAssay(Root_Integrated) <- "integrated"
Root_Integrated <- ScaleData(Root_Integrated, verbose = F)
```

# PCA
```{r}
Root_Integrated$Run <- Root_Integrated@meta.data$orig.ident

Root_Integrated <- RunPCA(Root_Integrated, 
                          features = VariableFeatures(object = Root_Integrated),
                          verbose = F)

DimPlot(Root_Integrated, reduction = "pca", cols = brewer.pal(8, "Set1")[1:2], 
        split.by = "Run", group.by = "Run")
```
```{r}
ElbowPlot(Root_Integrated)
```
# UMAP 
```{r}
Root_Integrated <- FindNeighbors(Root_Integrated, dims = 1:20)
Root_Integrated <- FindClusters(Root_Integrated, resolution = 0.5)
Root_Integrated <- RunUMAP(Root_Integrated, dims = 1:20,
                           min.dist = 0.001, repulsion.strength = 1, 
                           n.neighbors = 15, spread = 5)

DimPlot(Root_Integrated, reduction = "umap", label = T, label.size = 5, repel = T) +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) 
```

```{r}
UMAP_by_run <- DimPlot(Root_Integrated, reduction = "umap", label = T, label.size = 5, repel = T,
         split.by = "Run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by replicates\n")

UMAP_by_run
```
# Find cell type marker 
```{r}
root_10X_markers <- FindAllMarkers(Root_Integrated, only.pos = TRUE, 
                           min.pct = 0.25, logfc.threshold = 0.25, verbose = F) 

root_10X_markers <- root_10X_markers %>%  
  filter(p_val_adj < 0.01) 

head(root_10X_markers)
```


```{r}
root_10X_markers_top2 <- root_10X_markers %>% 
  group_by(cluster) %>% 
  slice_max(n = 5, order_by = avg_log2FC)


RootDotPlot1 <- DotPlot(Root_Integrated, features = root_10X_markers_top2$gene, 
        ) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Cluster",
       x = "Genes") + 
  coord_flip()

RootDotPlot1
```
```{r}
functional_annotation <- read_delim("../Data/cro_v3_anno/cro_v3.functional_annotation.txt",
                                    delim = "\t", col_names = F)

head(functional_annotation)
```
```{r}
cell_maker_functional <- functional_annotation %>% 
  mutate(gene = str_replace(X1, "_", "-")) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(root_10X_markers_top2, by = "gene") %>% 
  distinct(gene, .keep_all = T) %>% 
  arrange(cluster)
```


# Identify cell types
```{r}
root_markers <- read_csv("../Results/R_outputs/root_markers.csv")
root_markers <- root_markers %>% 
  mutate(gene = str_replace(cro.v3, "_", "-"))

head(root_markers)
```

```{r}
root_dotplot <- DotPlot(Root_Integrated, features = unique(root_markers$gene))
```

```{r}
root_dotplot$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  ggplot(aes(y = gene, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.8) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Average Exp.",
       size = "% Expressed") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black", face = "bold"),
        axis.text = element_text(color = "black"))
```
## Lignin biosynthesis 
```{r}
Orthogroups <- read_delim("../Results/OF_Results_Apr29/Orthologues/Orthologues_Araport_pep_20220119_representative_gene_model/Araport_pep_20220119_representative_gene_model__v__cro_v3.gene_models.repr.pep.tsv", 
                          delim = "\t", col_names = T, col_types = cols())

Orthogroups <- Orthogroups %>% 
  rename(
    Ath = Araport_pep_20220119_representative_gene_model, 
    Cro = cro_v3.gene_models.repr.pep
  )
head(Orthogroups)
```

```{r}
DotPlot(Root_Integrated, features = c("CRO-06G019550", # PAL
                                      "CRO-03G000880", # PAL
                                      "CRO-06G016650", # C4H 
                                      "CRO-08G032860", # 4CL
                                      "CRO-07G000130" # C3H 
                                      ) # 
        )+    
  coord_flip()
```


* 0: ?
* 1: meristem
* 2: root cap
* 3: ?
* 4: Endodermis/cortex 
* 5: Stele 
* 6: Stele/xylem
* 7: ?
* 8: epidermis atrichoblast
* 9: Stele

```{r}
root_cell_type_assignment <- data.frame(
  cluster = c(0:9)
) %>% 
  mutate(cell_type = case_when(
    cluster == 1 ~ "Meristem",
    cluster == 2 ~ "Root cap",
    cluster == 4 ~ "Ground",
    cluster == 5| 
      cluster == 9 ~ "Stele",
    cluster == 6 ~ "Stele", 
    cluster == 8 ~ "Epidermis",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, levels = c(
    "Meristem", "Root cap", "Ground", "Epidermis", "Stele", "Unassigned"
  ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "1", "2", "4", "8", "5", "6", "9", "0", "3", "7"
  )))

levels(Root_Integrated) <- c(
     "1", "2", "4", "8", "5", "6", "9", "0", "3", "7"
  )
```

```{r}
root_marker_nicer <- read_excel("../Data/cell_type_marker_root.xlsx") 
root_marker_nicer <- root_marker_nicer %>% 
  mutate(gene = str_replace_all(Cro_v3, "_", "-")) %>% 
  mutate(order = 1:nrow(root_marker_nicer))

head(root_marker_nicer) 

Root_marker_Dot <- DotPlot(Root_Integrated, features = rev(root_marker_nicer$gene))
```
```{r}
Root_markers_heatmap <- Root_marker_Dot$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(root_marker_nicer, by = "gene") %>% 
  mutate(gene = reorder(gene, -order)) %>% 
  ggplot(aes(y = gene, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.8) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg. Exp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "top",
        legend.box = "vertical")
```


```{r}
root_cell_type_strip <- root_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(5, "Set2"), "grey80")) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

root_cell_type_mkr_graph <- root_marker_nicer %>% 
  ggplot(aes(x = "", y = -order)) +
  geom_text(aes(label = Symbol)) +
  theme_void()
```

```{r}
wrap_plots(Root_markers_heatmap, root_cell_type_mkr_graph,
          root_cell_type_strip, 
          heights = c(1, 0.05), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)

ggsave("../Results/R_output_Root_scRNAseq/Cell_type_assignment_plot_root.svg",
       height = 6, width = 7, bg = "white")
ggsave("../Results/R_output_Root_scRNAseq/Cell_type_assignment_plot_root.png", 
       height = 6, width = 7, bg = "white")
```

### Enhanced UMAP 
```{r}
UMAP_root <- DimPlot(Root_Integrated, reduction = "umap", label = T, label.size = 5, repel = T) +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) 

root_cell_type_garph <- root_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "")) +
  facet_wrap(~ cell_type, scales = "free", ncol = 2) +
  geom_point(size = 5, aes(color = cluster)) +
  geom_text(aes(label = cluster)) +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black", face = "bold"),
    axis.text.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )

wrap_plots(UMAP_root, root_cell_type_garph, ncol = 1, heights = c(1, 1))

ggsave("../Results/R_output_Root_scRNAseq/UMAP_root.svg", 
       height = 3.25, width = 2.8, bg = "white")

ggsave("../Results/R_output_Root_scRNAseq/UMAP_root.png", 
       height = 3.25, width = 2.8, bg = "white")
```
```{r}
write_excel_csv(UMAP_root$data, file = "../../CRO_manuscript_figures/Source_Files/Fig7a_1.csv")
write_excel_csv(root_cell_type_garph$data, file = "../../CRO_manuscript_figures/Source_Files/Fig7a_2.csv")
```

```{r}
saveRDS(Root_Integrated,file = "../Results/R_output_Root_scRNAseq/Root_Integrated.Rds")
```


## UMAP-MAT
```{r}
UMAP_MAT <- FeaturePlot(Root_Integrated, features = "CRO-02G001000", 
                          order = T) +
  scale_color_gradientn(colors = brewer.pal(9, "PuBuGn")[1:9]) +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("MAT: clusters 4 & 8")

wrap_plots(UMAP_by_run, UMAP_MAT, nrow = 1, widths = c(1, 0.6))

ggsave("../Results/R_output_Root_scRNAseq/UMAP_MAT.svg", height = 3, width = 8)
ggsave("../Results/R_output_Root_scRNAseq/UMAP_MAT.png", height = 3, width = 8)
```
 
# Check pathway genes


## at cluster level 
```{r}
MIA_genes_info <- readRDS(file = "../Results/R_outputs/MIA_genes_info.Rds")

root_MIA_genes <- MIA_genes_info %>% 
  mutate(tag = case_when(
    str_detect(as.character(gene), "05G028810") ~ "7DLGT",
    str_detect(as.character(gene), "04G032090") ~ "GPPS(LSU)",
    T ~ tag
    )) %>% 
  filter(is.na(tag) == F) %>% 
  mutate(order = case_when(
    str_detect(as.character(gene), "04G032090") ~  9,
    str_detect(as.character(gene), "05G028810") ~ 15, 
    T ~ order
  )) %>% 
  arrange(order) %>% 
  mutate(segment = case_when(
    str_detect(as.character(gene), "04G032090") ~ "MEP",
    str_detect(as.character(gene), "05G028810") ~ "Iridoid", 
    T ~ segment
  )) %>% 
  #filter(is.na(order) == F) %>% 
  filter(str_detect(as.character(gene), "05G025350") == F) %>% 
  filter(str_detect(as.character(gene), "05G028780") == F) %>% 
  filter(str_detect(tag, "ADH|THAS|HYS|ISP|GS2|D4H|DAT|SS") == F) %>% 
   filter(str_detect(tag, "T3O|T3R|NMT|16O|T16H") == F) %>% 
  mutate(tag = case_when(
    str_detect(tag, "SGD$") ~ "SGD1",
    str_detect(tag, "MATE") ~ "SLTr",
    #str_detect(tag, "SGD2") ~ "SDG2",
    str_detect(tag, "GS$") ~ "GS1",
    T ~ tag
  )) %>% 
  mutate(gene = str_replace_all(gene, "_", "-"))  %>% 
  filter(segment != "TF") %>% 
  mutate(order = case_when(
    str_detect(tag, "TEX1") ~  38,
    str_detect(tag, "TEX2") ~  39,
    str_detect(tag, "T19H") ~  40,
    str_detect(tag, "TAT") ~  41,
    str_detect(tag, "MAT$") ~  42,
    T ~ order
  )) 

head(root_MIA_genes)
```

```{r}
MIA_at_cluster_root <- DotPlot(Root_Integrated, features = rev(root_MIA_genes$gene)) 
```

```{r}
MIA_at_cluster_root$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(root_MIA_genes, by = "gene") %>%
  mutate(gene = reorder(gene, -order)) %>%
  mutate(tag = reorder(tag, -order)) %>%
  ggplot(aes(y = tag, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.8) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Average Exp.",
       size = "% Expressed") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black", face = "bold"),
        axis.text = element_text(color = "black"))
```

## at cell type level 
```{r}
metadata_root <- Root_Integrated@meta.data %>% 
  mutate(barcode = row.names(Root_Integrated@meta.data)) %>% 
    mutate(cell_type = case_when(
    seurat_clusters == 1 ~ "Meristem",
    seurat_clusters == 2 ~ "Root cap",
    seurat_clusters == 4 ~ "Ground",
    seurat_clusters == 5| 
      seurat_clusters == 9 ~ "Stele",
    seurat_clusters == 6 ~ "Stele", 
    seurat_clusters == 8 ~ "Epidermis",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, levels = c(
    "Meristem", "Root cap", "Ground", "Epidermis", "Stele", "Unassigned"
  ))) 

head(metadata_root)

Root_Integrated$cell_type <- metadata_root$cell_type
```

```{r}
DotPlot_MIA_root_cell_type <- DotPlot(Root_Integrated, 
                                 features = rev(root_MIA_genes$gene),
                                 group.by = "cell_type")
```

```{r}
DotPlot_MIA_root_cell_type$data %>%
  mutate(id2 = case_when(
    str_detect(id, "Meristem") ~ "M",
    str_detect(id, "Root cap") ~ "RC",
    str_detect(id, "Epide") ~ "E",
    str_detect(id, "Stele") ~ "St",
    str_detect(id, "Ground") ~ "G",
    T ~ "Un"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "M", "RC", "G",
    "E", "St", "Un"
  ))) %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(root_MIA_genes, by = "gene") %>%
  mutate(gene = reorder(gene, -order)) %>%
  mutate(tag = reorder(tag, -order)) %>%
  ggplot(aes(y = tag, x = id2)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  # scale_y_discrete(label = NULL) +
  scale_fill_viridis(option = "A", begin = 0, end = 0.8) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"),
                             title.position = "top", nrow = 4)) +
  guides(fill = guide_colorbar(title.position = "top")) +
  labs(x = NULL,
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  #theme_minimal() +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y = element_text(hjust = 0.5),
        axis.text.x = element_text(hjust = 1, angle = 45, face = "plain"),
        legend.position = "right",
        legend.box = "vertical",
        legend.key.height = unit(0.8, "lines"),
        legend.key.width = unit(0.8, "lines"),
        legend.title = element_text(size = 12),
        title = element_text(size = 10, face = "bold")) 

ggsave("../Results/R_output_root_scRNAseq/MIA_at_cell_type_root.svg", height = 7, width = 3, bg = "white")
ggsave("../Results/R_output_root_scRNAseq/MIA_at_cell_type_root.png", height = 7, width = 3, bg = "white")
```

```{r}
write_excel_csv(
  DotPlot_MIA_root_cell_type$data %>%
  mutate(id2 = case_when(
    str_detect(id, "Meristem") ~ "M",
    str_detect(id, "Root cap") ~ "RC",
    str_detect(id, "Epide") ~ "E",
    str_detect(id, "Stele") ~ "St",
    str_detect(id, "Ground") ~ "G",
    T ~ "Un"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "M", "RC", "G",
    "E", "St", "Un"
  ))) %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(root_MIA_genes, by = "gene") %>%
  mutate(gene = reorder(gene, -order)) %>%
  mutate(tag = reorder(tag, -order)),
  file = "../../CRO_manuscript_figures/Source_Files/Fig7b.csv"
)
```





## UMAP colored by biosynthetic genes 

ISY, IO,
SRT, GS1,
TS, TAT
```{r}
root_MIA_genes
```

```{r}
Root_DXS <- FeaturePlot(Root_Integrated, features = c("CRO-02G005470"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("DXS\n02G005470") 

Root_ISY <- FeaturePlot(Root_Integrated, features = c("CRO-07G007680"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("ISY\n07G007680") 

#Root_ISY

Root_IO <- FeaturePlot(Root_Integrated, features = c("CRO-03G004950"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("IO\n03G004950") 

#Root_IO
```

```{r}
Root_TDC <- FeaturePlot(Root_Integrated, features = c("CRO-03G032340"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("TDC\n03G032340") 

Root_STR <- FeaturePlot(Root_Integrated, features = c("CRO-03G032330"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("STR\n03G032330") 

Root_GS <- FeaturePlot(Root_Integrated, features = c("CRO-06G024600"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("GS\n06G024600") 

#Root_TDC
```


```{r}
Root_TEX2 <- FeaturePlot(Root_Integrated, features = "CRO-05G001500", 
            cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("TEX2\n05G001500") 


Root_T19H <- FeaturePlot(Root_Integrated, features = "CRO-05G032890", 
            cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("T19H\n05G032890") 

Root_TAT <- FeaturePlot(Root_Integrated, features = c("CRO-02G001020"),
            cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("TAT\n02G001020")  
```

```{r}
wrap_plots(
  Root_DXS, Root_ISY , Root_IO,
  Root_TDC, Root_STR, Root_GS,
  Root_TEX2, Root_T19H, Root_TAT,
  ncol = 3, nrow = 3,
  guides = "collect"
) &
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 10)
    )

ggsave("../Results/R_output_Root_scRNAseq/FeaturePlot_root.svg", 
       height = 6, width = 5, bg = "white")

ggsave("../Results/R_output_Root_scRNAseq/FeaturePlot_root.png", 
       height = 6, width = 5, bg = "white")
```

```{r}
FeaturePlot(Root_Integrated, features = c("CRO-07G007670"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("ISY (07G007670)") 

FeaturePlot(Root_Integrated, features = c("CRO-07G007680"),
                      cols = brewer.pal(9, "PuBuGn"), shape.by = "Run", order = T, pt.size = 1) +
  guides(shape = guide_legend(override.aes = list(fill = "grey40", size = 3))) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black")
  ) +
  ggtitle("ISY (07G007680)") 
```


# Paralogs 
```{r}
paralogs <- readRDS("../Results/R_outputs/paralogs.Rds")

paralogs_nice <- paralogs %>% 
  mutate(gene = str_replace_all(v3, "_", "-"))

head(paralogs_nice)
```
## Root data
```{r}
Paralog_root <- DotPlot(Root_Integrated, features = rev(paralogs_nice$gene), 
                        group.by = "cell_type")
```

### Fill in missing values 
CRO-03G003550 -  SS
CRO-01G033090 -  THAS3 
CRO-03G003540 -  SS 

These 3 are not detected in root at all. 
```{r}
missing_in_root <- c("CRO-01G033090")
id2 <- c(
    "M", "RC", "G",
    "E", "St", "Un"
  ) 

missing_data_root <- expand.grid(
  gene = missing_in_root, 
  id2 = id2) %>% 
  mutate(tag = case_when(
    str_detect(gene, "CRO-01G033090") ~ "THAS3"
  )) %>% 
  mutate(avg.exp.scaled = NA) %>% 
  mutate(pct.exp = NA)

head(missing_data_root)
```

```{r}
Paralog_root_sc <- Paralog_root$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  full_join(paralogs_nice, by = "gene") %>% 
  mutate(tag = factor(tag, levels = c(
    "7DLGT", "ISY", "TEX2", "THAS3" 
  ))) %>%
  mutate(id2 = case_when(
    str_detect(id, "Meristem") ~ "M",
    str_detect(id, "Root cap") ~ "RC",
    str_detect(id, "Epide") ~ "E",
    str_detect(id, "Stele") ~ "St",
    str_detect(id, "Ground") ~ "G",
    T ~ "Un"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "M", "RC", "G",
    "E", "St", "Un"
  ))) %>% 
  select(avg.exp.scaled, pct.exp, gene, tag, id2) %>% 
  filter(is.na(id2) == F) %>% 
  rbind(
    missing_data_root
  ) %>% 
  filter(str_detect(gene, "07G007660|07G007690") == F) %>% 
  ggplot(aes(y = gene, x = id2)) +
  facet_grid(tag ~ ., scales = "free_y", space = "free_y", switch = "y") +
  geom_point(aes(fill = avg.exp.scaled, size = pct.exp), shape = 21, color = "white") +
  #geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"), na.value = "grey80",
                       breaks = c(0, 1)) +
  scale_size(breaks = c(40, 60)) +
  scale_y_discrete(label = NULL) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), order = 1, nrow = 2)) +
  labs(x = NULL,
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.key.width = unit(0.8, "lines"),
    legend.title = element_text(size = 12, hjust = 0.5),
    legend.justification = 1,
    legend.box.margin = margin(t = -1, unit = ("lines")),
    legend.box = "vertical",
    panel.border = element_rect(color = "grey20", fill =NA),
    plot.title = element_text(size = 10)
 ) +
  ggtitle("root scRNA-seq")

Paralog_root_sc
```
## Leaf data 
```{r}
Paralog_leaf_sc <- readRDS(file = "../Results/R_output_Leaf_scRNAseq/Paralog_leaf_sc.Rds")
Paralog_leaf_sc

Paralog_legend <- get_legend(Paralog_leaf_sc)
```

## combine together 
```{r}
Paralog_dots <- wrap_plots(Paralog_leaf_sc,
           Paralog_root_sc, 
           ncol = 2) &
  theme(legend.position = "none")

plot_grid(Paralog_dots, Paralog_legend, ncol = 1, 
          rel_heights = c(1, 0.1))

ggsave("../Results/R_output_Root_scRNAseq/Paralogs_combined.svg", 
       height = 4, width = 3.5, bg = "white")

ggsave("../Results/R_output_Root_scRNAseq/Paralogs_combined.png", 
       height = 4, width = 3.5, bg = "white")
```
```{r}
write_excel_csv(Paralog_leaf_sc$data, "../../CRO_manuscript_figures/Source_Files/Fig7c_1.csv")
write_excel_csv(Paralog_root_sc$data, "../../CRO_manuscript_figures/Source_Files/Fig7c_2.csv")
```

# Peak expression for pathway genes 
```{r}
expr.root <- GetAssayData(Root_Integrated, slot = "scale.data")

expr.root_by_cell_type <- expr.root %>% 
  as.data.frame() %>% 
  mutate(gene = row.names(.)) %>% 
  gather("barcode", "exp", 1:nrow(metadata_root)) %>%
  inner_join(metadata_root, by = "barcode") %>% 
  group_by(gene, cell_type) %>% 
  summarise(mean.rel.exp = mean(exp)) %>% 
  ungroup()

head(expr.root_by_cell_type)
```

```{r}
expr.root_by_cell_type_max <- expr.root_by_cell_type %>% 
  group_by(gene) %>% 
  slice_max(order_by = mean.rel.exp, n = 1, with_ties = F) %>% 
  ungroup()

head(expr.root_by_cell_type_max)

write_excel_csv(expr.root_by_cell_type_max, "../Results/R_output_Root_scRNAseq/Root_max_exp.csv")
```


# For Chloe
## Data
```{r}
Chloe_gene_list <- read_excel("../Data/From_Chloe/gene_coexpression.xlsx")

Chloe_gene_list <- Chloe_gene_list %>% 
  mutate(gene = str_replace(Contig, "_", "-"))

head(Chloe_gene_list)
```
```{r}
metadata_root2 <- Root_Integrated@meta.data %>% 
  mutate(barcode = row.names(Root_Integrated@meta.data)) %>% 
    mutate(cell_type = case_when(
    seurat_clusters == 1 ~ "Meristem",
    seurat_clusters == 2 ~ "Root cap",
    seurat_clusters == 4 ~ "Ground",
    seurat_clusters == 5| 
      seurat_clusters == 9 ~ "Stele",
    seurat_clusters == 6 ~ "Stele", 
    seurat_clusters == 8 ~ "Epidermis",
    seurat_clusters == 0 ~ "Unassigned 1",
    seurat_clusters == 3 ~ "Unassigned 2",
    seurat_clusters == 7 ~ "Unassigned 3",
  )) %>% 
  mutate(cell_type = factor(cell_type, levels = c(
    "Meristem", "Root cap", "Ground", "Epidermis", "Stele", 
    "Unassigned 1", "Unassigned 2", "Unassigned 3"
  ))) 

head(metadata_root2)

Root_Integrated$cell_type2 <- metadata_root2$cell_type
```

## Graph
```{r}
Chloe_Dopotplot_root <- DotPlot(Root_Integrated, features = Chloe_gene_list$gene, group.by = "cell_type2") 
```

```{r}
Chloe_Dopotplot_root$data %>%  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  mutate(id2 = case_when(
    str_detect(id, "Meristem") ~ "M",
    str_detect(id, "Root cap") ~ "RC",
    str_detect(id, "Epide") ~ "E",
    str_detect(id, "Stele") ~ "St",
    str_detect(id, "Ground") ~ "G",
    str_detect(id, "Unassigned 1") ~ "Un 1",
    str_detect(id, "Unassigned 2") ~ "Un 2",
    str_detect(id, "Unassigned 3") ~ "Un 3"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "M", "RC", "G",
    "E", "St", 
    "Un 1", "Un 2", "Un 3"
  ))) %>% 
  inner_join(Chloe_gene_list, by = "gene") %>% 
  mutate(Gene_name = reorder(`Gene name`, -Order)) %>% 
  mutate(Pathway = factor(Pathway, levels = c(
    "MIA", "Flavonoid", "Phenylpropanoid", "Lignin"
  ))) %>% 
  ggplot(aes(y = Gene_name, x = id2)) +
  facet_grid(Pathway ~., scales = "free_y", space = "free_y", switch = "y") +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", end = 0.9, breaks = c(0, 2)) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), 
                             order = 2, ncol = 1)) +
  labs(x = "Cell types",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(hjust = 0.5),
    legend.position = "right",
    legend.box = "vertical", 
    legend.key.width = unit(0.8, "lines"),
    legend.title = element_text(size = 12, hjust = 0.5),
    panel.spacing = unit(1, "line"),
    strip.background = element_blank(),
    strip.placement = "outside",
    panel.border = element_rect(fill = NA),
    strip.text = element_blank() 
 )

ggsave("../Results/R_output_Root_scRNAseq/Chloe_genes_root.svg", height = 7, width = 3.5, bg = "white")
ggsave("../Results/R_output_Root_scRNAseq/Chloe_genes_root.png", height = 7, width = 3.5, bg = "white")
```
## Chr 8 end
```{r}
Chloe_Chr8 <- read_excel("../../CRO_v3/Data/From_Chloe/chr8_cluster.xlsx", col_names = F)
colnames(Chloe_Chr8) <- c("gene", "annotation", "tag")

Chloe_Chr8 <- Chloe_Chr8 %>% 
  distinct(gene, .keep_all = T) %>% 
  mutate(gene2 = str_replace(gene, "_", "-")) %>% 
  mutate(order = 1:nrow(.))

head(Chloe_Chr8)
```

```{r}
Chloe_Chr8_Dopotplot_root <- DotPlot(Root_Integrated, features = Chloe_Chr8$gene2, group.by = "cell_type2") 
```
```{r}
Dotplot_Chr8_root <- Chloe_Chr8_Dopotplot_root$data %>%  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  #filter(str_detect(internal.tag, "rna") == F) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  mutate(id2 = case_when(
    str_detect(id, "Meristem") ~ "M",
    str_detect(id, "Root cap") ~ "RC",
    str_detect(id, "Epide") ~ "E",
    str_detect(id, "Stele") ~ "St",
    str_detect(id, "Ground") ~ "G",
    str_detect(id, "Unassigned 1") ~ "Un 1",
    str_detect(id, "Unassigned 2") ~ "Un 2",
    str_detect(id, "Unassigned 3") ~ "Un 3"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "M", "RC", "G",
    "E", "St", 
    "Un 1", "Un 2", "Un 3"
  ))) %>% 
  inner_join(Chloe_Chr8, by = c("gene"="gene2")) %>% 
  mutate(gene_id = reorder(gene, -order)) %>% 
  ggplot(aes(y = gene_id, x = id2)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", end = 0.9, breaks = c(0, 2)) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), 
                             order = 2, ncol = 1)) +
  labs(x = "Cell types",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(hjust = 0.5),
    legend.position = "right",
    legend.box = "vertical", 
    legend.key.width = unit(0.8, "lines"),
    legend.title = element_text(size = 12, hjust = 0.5)
 )

Dotplot_Chr8_root

ggsave("../Results/R_output_Root_scRNAseq/Chloe_Chr8_end_root.svg", height = 7.2, width = 4, bg = "white")
ggsave("../Results/R_output_Root_scRNAseq/Chloe_Chr8_end_root.png", height = 7.2, width = 4, bg = "white")
```

