---
title: "SingleCellMetabolites"
author: "Chenxin Li"
date: "6/30/2022"
output:  
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages 
```{r}
library(tidyverse)
library(umap)
library(dynamicTreeCut)
library(igraph)
library(ggraph)
library(readxl)
library(RColorBrewer)
library(viridis)
library(svglite) 
library(patchwork)
library(cowplot)
```


# Data 
```{r}
Corrected_Areas_clean <- read_csv("../Data/Last_attempt/Corrected_Areas_clean_1e5_last.csv")

Corrected_Areas_clean <- Corrected_Areas_clean %>% 
  rename(feature = `...1`)

head(Corrected_Areas_clean)
```

```{r}
Corrected_Areas_clean %>% 
  filter(str_detect(feature, "[a-z]"))
```

# Log Normalize and assign plate ID 
```{r}
Area_long <- Corrected_Areas_clean %>% 
  gather("cell", "area", 2:ncol(Corrected_Areas_clean)) %>% 
  filter(str_detect(cell, "QC") == F) %>% 
  mutate(log_Area = log10(area + 1)) %>% 
  mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  ))

head(Area_long)
```

```{r}
Area_long %>% 
  group_by(feature, Plate) %>% 
  filter(feature == "Secologanin") %>% 
  count() %>% 
  arrange(Plate)  %>% 
  ungroup() 

Area_long %>% 
  group_by(feature, Plate) %>% 
  filter(feature == "Secologanin") %>% 
  count() %>% 
  arrange(Plate)  %>% 
  ungroup() %>% 
  summarise(total = sum(n))
```

552 total cells 


## Check distribution 
```{r}
Area_long_s <- Area_long %>% 
  filter(area > 0) %>%
  group_by(Plate) %>% 
  summarise(
    mean = mean(log_Area),
    sd = sd(log_Area)
  )

Area_long_s
```

```{r}
Area_long %>% 
  filter(area > 0) %>% 
  ggplot(aes(x = log_Area)) +
  facet_grid(Plate ~., scales = "free_y") +
  geom_histogram(bins = 100, alpha = 0.8, color = "white", fill = "grey20") +
  geom_vline(data = Area_long_s, aes(xintercept = mean), 
             size = 1, color = "tomato1", alpha = 0.8) +
  labs(x = "log10 Area",
       y = "non-zero occurrences") +
  theme_classic() +
  theme(
    panel.spacing = unit(1, "lines")
  )

ggsave("../Results/R_outputs_multiplate/hist_by_batch_5e4.svg", height = 7, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/hist_by_batch_5e4.png", height = 7, width = 5, bg = "white")
```


```{r}
Area_long %>% 
  group_by(Plate) %>% 
  filter(area == 0) %>% 
  count()
```

# Without batch correction
## Standardize together  
```{r}
Area_long <- Area_long %>% 
  group_by(feature) %>% 
  mutate(z = (log_Area - mean(log_Area))/sd(log_Area)) %>% 
  filter(is.na(z) == F) %>% 
  ungroup()

head(Area_long)
```
## PCA 
```{r}
Area_z_wide <- Area_long %>% 
  select(feature, cell, z) %>%
  spread(cell, z) %>% 
  as.data.frame()

row.names(Area_z_wide) <- Area_z_wide$feature

head(Area_z_wide)
```
```{r}
pc <- prcomp(t(Area_z_wide[, -1])) 
```

```{r}
pc$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = Plate)) +
  theme_bw()

ggsave("../Results/R_outputs_multiplate/PCA1_5e4.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/PCA1_5e4.png", height = 4, width = 5, bg = "white")
```

# Batch correction
## select features
```{r}
Common_features <- Area_long %>% 
  group_by(feature, Plate) %>% 
  summarise(mean.area = mean(area)) %>% 
  filter(mean.area > 0) %>% 
  count() %>% 
  filter(n >= 7)

nrow(Common_features)
head(Common_features)
```

Dimension reduction using 932 commonly detected features 

```{r}
feature_detect_bar <- Area_long %>% 
  group_by(feature, Plate) %>% 
  summarise(mean.area = mean(area)) %>% 
  filter(mean.area > 0) %>% 
  count() %>% 
  rename(detected_in = n) %>% 
  ungroup() %>% 
  group_by(detected_in) %>% 
  count() %>% 
  ggplot(aes(x = detected_in, y = n)) +
  geom_bar(stat = "identity", alpha = 0.8, width = 0.7) +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(x = "Detected in how many batches",
       y = "No. of features") +
  theme_classic()

feature_detect_bar
```



```{r}
feature_detect_bee <- Area_long %>% 
  filter(area > 0) %>% 
  group_by(feature) %>% 
  summarise(mean.log.area = mean(log_Area)) %>% 
  inner_join(
    Area_long %>% 
      group_by(feature, Plate) %>% 
      summarise(mean.area = mean(area)) %>% 
      filter(mean.area > 0) %>% 
      count(), 
    by = "feature"
  ) %>% 
  ggplot(aes(x = n , y = mean.log.area)) +
  ggbeeswarm::geom_quasirandom(alpha = 0.8) +
  stat_summary(geom = "point", fun = mean, color = "tomato1", size = 3) +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(x = NULL,
       y = "log10 peak area\n(only non zero peaks are shown)",
       title = "Red = mean of non-zero peaks") +
  theme_classic() +
  theme(plot.title = element_text(size = 10))

feature_detect_bee 

wrap_plots(feature_detect_bee, feature_detect_bar, nrow = 2)

ggsave("../Results/R_outputs_multiplate/feature_detection_5e4.svg", height = 5, width = 3, bg = "white")
ggsave("../Results/R_outputs_multiplate/feature_destection_5e4.png", height = 5, width = 3, bg = "white")
```

Commonly detected features also have the highest mean log area (for non-zero values). 

## Standardize by feature and batch
```{r}
Area_long_batch <- Area_long %>% 
  filter(feature %in% Common_features$feature) %>%  
  # select only features commonly detected in all batches 
  group_by(feature, Plate) %>% 
  mutate(z = (log_Area - mean(log_Area))/sd(log_Area)) %>% 
  ungroup()

head(Area_long_batch)
```


## Check distribution again 
```{r}
Area_long_batch_s <- Area_long_batch %>% 
  filter(area > 0) %>%
  group_by(Plate) %>% 
  summarise(
    mean.z = mean(z),
    sd.z = sd(z)
  )

Area_long_batch_s
```

```{r}
Area_long_batch %>% 
  filter(area > 0) %>% 
  ggplot(aes(x = z)) +
  facet_grid(Plate ~., scales = "free_y") +
  geom_histogram(bins = 100, alpha = 0.8, color = "white", fill = "grey20") +
  #geom_vline(data = Area_long_s, aes(xintercept = mean), 
   #          size = 1, color = "tomato1", alpha = 0.8) +
  labs(x = "z score",
       y = "non-zero occurrences") +
  theme_classic() +
  theme(
    panel.spacing = unit(1, "lines")
  )
```

## PCA
```{r}
Area_z_wide_batch <- Area_long_batch %>% 
  select(feature, cell, z) %>%
  spread(cell, z) %>% 
  as.data.frame()

row.names(Area_z_wide_batch) <- Area_z_wide_batch$feature

head(Area_z_wide_batch)
```

```{r}
pc_batch <- prcomp(t(Area_z_wide_batch[, -1])) 
```

```{r}
pc_batch$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = Plate)) +
  theme_bw()

ggsave("../Results/R_outputs_multiplate/PCA_batch_5e4.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/PCA_batch_5e4.png", height = 4, width = 5, bg = "white")
```
```{r}
pc_batch$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  inner_join(
   Area_long %>% 
   filter(feature == "Serpentine") %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(fill = z), shape = 21, alpha = 0.8 , size = 3) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  theme_bw() +
  ggtitle("Serpentine") 
```

### Correlate each feature to each PC
```{r}
feature_by_cell <- Area_long_batch %>% 
  select(cell, feature, z) %>% 
  spread(feature, z) %>% 
  as.data.frame()

row.names(feature_by_cell) <- feature_by_cell$cell

head(feature_by_cell)

#head(pc_batch$x)
```

```{r}
feature_by_cell %>% 
  ggplot(aes(x = Serpentine, y = Secologanin)) +
  geom_point(alpha = 0.5) +
  labs(x = "Serpentine z score",
       y = "Secologanin z score") +
  theme_classic()
```


```{r}
feature_PC_cor <- function(feature){
  cor(feature, pc_batch$x)
}
```

```{r}
feature_by_PC <- apply(X = feature_by_cell[, -1], FUN = feature_PC_cor, MARGIN = 2)
# slow step 
```

```{r}
feature_by_PC_df <- feature_by_PC %>% 
  as.data.frame() %>% 
  mutate(index = 1:nrow(feature_by_PC)) %>% 
  mutate(PC = paste0("PC", index)) %>% 
  gather("feature", "r", 1:nrow(Common_features))

head(feature_by_PC_df)
```

 


### Make heat map 
```{r}
feature_by_PC_df %>% 
  filter(index <= 50) %>% 
  filter(str_detect(feature, "[a-z]")) %>% 
  mutate(Rows = case_when(
    str_detect(feature, "Sec|Akua|Loga|Epi") ~ 1, 
    T ~ 2
  )) %>% 
  mutate(feature = reorder(feature, Rows)) %>% 
  #mutate(PC = reorder(PC, index)) %>% 
  ggplot(aes(x = index, y = feature)) +
  geom_tile(aes(fill = r)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")),
                       limits = c(-0.8, 0.8),
                       breaks = c(-0.5, 0, 0.5)) +
  labs(x = "PC",
       y = NULL) +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.y = element_text(hjust = 0.5)
  )

ggsave("../Results/R_outputs_multiplate/feature_vs_PC_5e4.svg", height = 5, width = 7, bg = "white")
ggsave("../Results/R_outputs_multiplate/feature_vs_PC_5e4.png", height = 5, width = 7, bg = "white")
```
I could have just project PC4 and PC11(?)  

```{r}
ordering_features <- feature_by_PC_df %>% 
  filter(index <= 50) %>% 
  group_by(feature) %>% 
  slice_max(n = 1, order_by = abs(r)) %>% 
  arrange(index) %>% 
  rename(max.index = index) %>% 
  ungroup() %>% 
  group_by(max.index) %>% 
  arrange(r, .by_group = T) %>% 
  #mutate(rank1 = rank(r, ties.method = "first")) %>% 
  mutate(order = max.index+1 - r) %>% 
  select(-PC, -r) %>% 
  ungroup()

ordering_features
```

```{r}
feature_by_PC_df %>% 
  filter(index <= 50) %>% 
  inner_join(ordering_features, by = "feature") %>% 
  mutate(feature = reorder(feature, -order)) %>% 
  arrange(r, .by_group = T) %>% 
  ggplot(aes(x = index, y = feature)) +
  geom_tile(aes(fill = r)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")),
                       limits = c(-0.9, 0.9),
                       breaks = c(-0.5, 0, 0.5)) +
  labs(x = "PC",
       y = NULL) +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.y = element_blank()
  )

ggsave("../Results/R_outputs_multiplate/feature_vs_PC_big_5e4.svg", height = 16, width = 16, bg = "white")
ggsave("../Results/R_outputs_multiplate/feature_vs_PC_big_5e4.png", height = 16, width = 16, bg = "white")
```


## Elbow plot
```{r}
pc_importance <- as.data.frame(t(summary(pc_batch)$importance))
pc_importance
```

```{r}
elbow <- pc_importance %>%
  mutate(PC = row.names(.)) %>% 
  mutate(PC = reorder(PC, -`Proportion of Variance`)) %>% 
  ggplot(aes(x = 1:nrow(pc_importance), y = `Proportion of Variance`)) +
  geom_point(size = 3, alpha = 0.8, shape = 21, 
             color = "white", fill = "grey20") +
  labs(x = "PC") +
  theme_classic()

cumu <- pc_importance %>%
  mutate(PC = row.names(.)) %>% 
  mutate(PC = reorder(PC, -`Proportion of Variance`)) %>% 
  ggplot(aes(x = 1:nrow(pc_importance), y = `Cumulative Proportion`)) +
  geom_point(size = 3, alpha = 0.8, shape = 21, 
             color = "white", fill = "grey20") +
  labs(x = "PC") +
  theme_classic()

wrap_plots(elbow, cumu, 
           design = "A
                     B")

ggsave("../Results/R_outputs_multiplate/PCA_initiation_5e4.svg", height = 6, width = 3)
ggsave("../Results/R_outputs_multiplate/PCA_initiation_5e4.png", height = 6, width = 3)
```
### Enchanced PCA plots 
```{r}
PCA_serp <- pc_batch$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  inner_join(
   Area_long_batch %>% 
   filter(feature == "Serpentine") %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  ggplot(aes(x = PC4, y = PC11)) +
  geom_point(aes(fill = z), shape = 21, alpha = 0.8 , size = 2) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 2)) +
  labs(y = "\nPC11") +
  theme_bw() +
  theme(plot.title = element_text(size = 12),
        legend.position = "bottom") +
  ggtitle("Serpentine")

PCA_serp

ggsave("../Results/R_outputs_multiplate/PCA_Serp_5e4.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/PCA_Serp_5e4.png", height = 4, width = 5, bg = "white")
```

```{r}
PCA_sec <- pc_batch$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  inner_join(
   Area_long_batch %>% 
   filter(feature == "Secologanin") %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  ggplot(aes(x = PC4, y = PC11)) +
  geom_point(aes(fill = z), shape = 21, alpha = 0.8 , size = 2) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  theme_bw() +
  theme(plot.title = element_text(size = 12),
        legend.position = "bottom") +
  ggtitle("Secologanin")

PCA_sec

ggsave("../Results/R_outputs_multiplate/PCA_Sec_5e4.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/PCA_Sec_5e4.png", height = 4, width = 5, bg = "white")
```



PC4 separates idioblast from everything  else

```{r}
wrap_plots(PCA_sec, PCA_serp,
           nrow = 1)

ggsave("../Results/R_outputs_multiplate/PCA_separation.svg", height = 3.5, width = 6, bg = "white")
ggsave("../Results/R_outputs_multiplate/PCA_separation.png", height = 3.5, width = 6, bg = "white")
```


# Supervised classification 
```{r}
pc_batch$x %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  inner_join(
   Area_long_batch %>% 
   #filter(feature == "Secologanin") %>% 
   filter(str_detect(feature, "Sec")) %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  ggplot(aes(x = PC27, y = z)) +
  geom_point() +
  theme_classic()
```
```{r}
cell_classification <- feature_by_cell %>% 
  #filter(str_detect(feature, "[a-z]")) %>% 
  #filter(feature %in% common_31$feature) %>% 
  mutate(cell_type = case_when(
    Secologanin > 1 & 
      Serpentine < 1 ~ "epidermis",
    Serpentine > 1 ~ "idioblast",
    T ~ "other"
  )) %>% 
    select(cell, cell_type)

head(cell_classification)
```
```{r}
Area_long_batch_cell_type <- Area_long %>% 
  full_join(cell_classification, by = "cell")  

head(Area_long_batch_cell_type)

Area_long_batch_cell_type %>% 
  filter(feature == "Secologanin") %>% 
  group_by(cell_type) %>% 
  count()
```
## plot
```{r}
 

Area_long_batch_cell_type %>% 
  filter(str_detect(feature, "[a-z]")) %>%
  filter(is.na(z) == F) %>% 
  mutate(feature = case_when(
    str_detect(feature, "_iminium") ~ "AHVB iminium",
    str_detect(feature, "Anhydrovinblastine$") ~ "AHVB",
    T ~ feature
  )) %>% 
  mutate(Rows = case_when(
    str_detect(feature, "Loganic") ~ 1,
    str_detect(feature, "Sec|Akua|Epi") ~ 2, 
    str_detect(feature, "Catha") ~ 3,
    str_detect(feature, "Vinblastine") ~ 5,
    T ~ 4
  )) %>% 
  mutate(z2 = case_when(
    z >= 3 ~ 3,
    T ~ z
  )) %>% 
  filter(str_detect(feature, "Unk_|acet") == F) %>% 
  mutate(feature = reorder(feature, -Rows)) %>% 
  mutate(cell_type_2 = case_when(
    str_detect(cell_type, "epi") ~ "E",
    str_detect(cell_type, "idio") ~ "I",
    T ~ "other"
  )) %>% 
  mutate(cell_type_2 = factor(cell_type_2, levels = c("E", "I", "other"))) %>% 
  ggplot(aes(x = cell, y = feature)) +
  facet_grid(.~cell_type_2, scales = "free_x", space = "free_x") +
  geom_tile(aes(fill = z2)) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"), 
                       breaks = c(0 ,3),
                       labels = c("0", ">3")
                       ) +
  scale_x_discrete(label = NULL) +
  labs(x = "Cell",
       y = "Compound",
       fill = "z score") +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text.y = element_text(color = "black", hjust = 0.5),
    axis.ticks.x = element_blank(),
    legend.position = "top"
  )

ggsave("../Results/R_outputs_multiplate/cell_heatmap_5e4.svg", height = 3.8, width = 5, bg = "white")
ggsave("../Results/R_outputs_multiplate/cell_heatmap_5e4.png", height = 3.8, width = 5, bg = "white")
```


```{r}
Area_long_batch_cell_type %>% 
  filter(str_detect(feature, "[a-z]")) %>% 
  group_by(feature) %>% 
  count()
```



# UMAP 
```{r}
custom.config <- umap.defaults
#custom.config$min_dist <- 0.5
#custom.config$spread <- 2
#custom.config$random_state <- 666
#custom.config$knn <- 10

run1.umap <- umap(pc_batch$x[, c(1:50)], config = custom.config)

head(run1.umap$layout)

```

## Additional meta data 
```{r}
Cat_cells <- read_csv("../Data/Single_Cell_metabolomics/Cat_cells.csv")

Cat_cells_long <- Cat_cells %>% 
  gather("cell", "area", 3:ncol(Cat_cells))
  

head(Cat_cells_long)
```
```{r}
suspected_idio <- read_excel("../Data/suspected_idioblast.xlsx")

suspected_idio <- suspected_idio %>% 
  mutate(cell = paste0(
    str_replace(Cell, "Cell", "C"), "_", Plate
  ))

head(suspected_idio)
```


## graph 
```{r}
UMAP_by_Plate <- run1.umap$layout %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(color = Plate)) +
  scale_color_manual(values = brewer.pal(7, "Set2")) +
  guides(color = guide_legend(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )

UMAP_by_Plate
```


```{r}
UMAP_by_cat <- run1.umap$layout %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  mutate(Cat = case_when(
    cell %in% Cat_cells_long$cell ~ "catharanthine",
    T ~ "no catharanthine"
  )) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(color = Cat)) +
  scale_color_manual(values = c("tomato1", "grey70")) +
  labs(color = NULL) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )

UMAP_by_cat
```

```{r}
UMAP_by_morph <- run1.umap$layout %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  mutate(Idioblast = case_when(
    cell %in% suspected_idio$cell ~ "likely idioblast",
    T ~ "others"
  )) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(color = Idioblast)) +
  scale_color_manual(values = c("slateblue3", "grey70")) +
  labs(color = NULL) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )

UMAP_by_morph
```
```{r}
wrap_plots(UMAP_by_Plate, UMAP_by_cat, UMAP_by_morph, nrow = 1)

ggsave("../Results/R_outputs_multiplate/UMAP_after_correction_5e4.png", width = 8, height = 4)
ggsave("../Results/R_outputs_multiplate/UMAP_after_correction_5e4.svg", width = 8, height = 4)
```
```{r}
UMAP_corrected <- run1.umap$layout %>% 
  as.data.frame() %>% 
  mutate(cell = row.names(.)) %>% 
   mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) 
```


# Confusion matrix for serpentine vs. idioblast suspects
```{r}
number_of_cells <- ncol(Area_z_wide_batch) - 1

Num.idioblast.suspect <-UMAP_corrected %>% 
  filter(cell %in% suspected_idio$cell) %>% 
  nrow()
 
Num.sep.positive <-UMAP_corrected %>% 
  inner_join(
   Area_long_batch %>% 
   filter(feature == "Serpentine") %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  filter(area > 1e4) %>% 
  nrow()
  
Num.sep.positive_idioblast.sus <- UMAP_corrected %>% 
  inner_join(
   Area_long_batch %>% 
   filter(feature == "Serpentine") %>% 
   arrange(z),
     by = c("cell", "Plate")
  ) %>% 
  filter(area > 0) %>% 
  filter(cell %in% suspected_idio$cell) %>% 
  nrow()

confusion_matrix <- matrix(c(Num.sep.positive_idioblast.sus, 
                             Num.idioblast.suspect - Num.sep.positive_idioblast.sus,
                             Num.sep.positive - Num.sep.positive_idioblast.sus,
                             number_of_cells - Num.idioblast.suspect - Num.sep.positive + Num.sep.positive_idioblast.sus), 
                           nrow = 2, ncol = 2) 

row.names(confusion_matrix) <- c("serpentine\ndetected", "serpentine\nnot detected")
colnames(confusion_matrix) <- c("idioblast\nmorphology", "not idioblast\nmorphology")

confusion_matrix

fisher.test(confusion_matrix)
```

## Plot confusion matrix
```{r}
confusion_matrix %>% 
  as.data.frame() %>% 
  mutate(serpentine = row.names(confusion_matrix)) %>% 
  gather("idioblast morphology", "count", 1:2) %>% 
  mutate(`idioblast morphology` = fct_rev(`idioblast morphology`)) %>% 
  ggplot(aes(y = `idioblast morphology`,
             x = `serpentine`)) +
  geom_tile(color = "black", size = 1, fill = NA) +
  geom_text(aes(label = count)) +
  scale_x_discrete(position = "top") +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(hjust = 0.5)
  )

ggsave("../Results/R_outputs_multiplate/idioblast_confusion_matrix_5e4.svg", 
       height = 2.5, width = 3, bg = "white")

ggsave("../Results/R_outputs_multiplate/idioblast_confusion_matrix_5e4.png", 
       height = 2.5, width = 3, bg = "white")
```


# Explore known compounds on UMAP 
## Idioblast  
```{r}
AHVB <- Area_long %>% 
  filter(feature == "Anhydrovinblastine") %>% 
  arrange(z) %>% 
  full_join(UMAP_corrected, by = c("cell", "Plate")) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(fill = z,), 
             alpha = 0.8, color = "grey20", shape = 21, size = 2) +
  #scale_fill_manual(values = brewer.pal(3, "Accent")) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  guides(fill = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12),
    panel.border = element_rect(fill = NA, color = "grey20")
  ) +
  ggtitle("Anhydrovinblastine")

AHVB
```

```{r}
SS <- Area_long %>% 
  filter(feature == "Serpentine") %>% 
  arrange(z) %>%
  full_join(UMAP_corrected, by = c("cell", "Plate")) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(fill = z), 
             alpha = 0.8, color = "grey20", shape = 21, size = 2) +
  #scale_fill_manual(values = brewer.pal(3, "Accent")) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  guides(fill = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12),
    panel.border = element_rect(fill = NA, color = "grey20")
  ) +
  ggtitle("Serpentine")

SS 
```

 

```{r}
Precon <- Area_long %>% 
  filter(str_detect(feature, "Preco")) %>% 
  arrange(z) %>% 
  full_join(UMAP_corrected, by = c("cell", "Plate")) %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(fill = z), 
             alpha = 0.8, color = "grey20", shape = 21, size = 2) +
  #scale_fill_manual(values = brewer.pal(3, "Accent")) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  guides(fill = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12),
    panel.border = element_rect(fill = NA, color = "grey20")
  ) +
  ggtitle("Precondylocarpine Acetate")

Precon
```

## Epidermis 
```{r}
Sec <- Area_long %>% 
  filter(feature == "Secologanin") %>% 
  full_join(UMAP_corrected, by = c("cell", "Plate")) %>% 
  arrange(z) %>%
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(fill = z), 
             alpha = 0.8, color = "grey20", shape = 21, size = 2) +
  #scale_fill_manual(values = brewer.pal(3, "Accent")) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"),
                       breaks = c(0, 1)) +
  guides(fill = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12),
    panel.border = element_rect(fill = NA, color = "grey20")
  ) +
  ggtitle("Secologanin")

Sec
```

# Write tables for LC
```{r}
Area_wide_batch <- Area_long_batch %>% 
  select(feature, cell, area) %>%
  spread(cell, area) %>% 
  as.data.frame()

row.names(Area_wide_batch) <- Area_wide_batch$feature

head(Area_wide_batch)
```


```{r}
# head(Area_z_wide_batch)
# head(Area_log_wide)

write_excel_csv(Area_z_wide_batch, "../Results/R_outputs_multiplate/z_scores_5e4.csv")
write_excel_csv(Area_wide_batch, "../Results/R_outputs_multiplate/peak_area_5e4.csv")
```


# graph based clustering
```{r}
pc.dist <- dist((pc_batch$x)[, 1:50], method = "e")
pc.sim <- (max(pc.dist) - as.matrix(pc.dist))/max(pc.dist) 

cellgraph <- graph_from_adjacency_matrix(pc.sim, mode = "upper",
                                         weighted = T, diag = F) 
```

```{r}
leiden_clusters <- cluster_leiden(cellgraph, resolution_parameter = 1, objective_function = "modularity")

leiden <- data.frame(
  cell = names(membership(leiden_clusters)),
  leiden = membership(leiden_clusters) %>% as.vector()
) %>% 
  mutate(cell = str_remove(cell, "_d1")) %>% 
  mutate(Plate = case_when(
   str_detect(cell, "p1d1") ~ "p1d1",
   str_detect(cell, "p1d2") ~ "p1d2",
   str_detect(cell, "p1d3") ~ "p1d3",
   str_detect(cell, "p2d1") ~ "p2d1",
   str_detect(cell, "p2d2") ~ "p2d2",
   str_detect(cell, "p2d3") ~ "p2d3",
   str_detect(cell, "p3d3") ~ "p3d3"
  )) %>% 
  mutate(cluster = as.factor(leiden))

head(leiden)
```
```{r}
leiden %>% 
  group_by(cluster) %>% 
  count()
```

```{r}
UMAP_corrected %>% 
  full_join(leiden, by = "cell") %>% 
  ggplot(aes(x = V1, y = V2)) +
  geom_point(aes(fill = cluster), 
             alpha = 0.8, color = "white", shape = 21, size = 3) +
  scale_fill_manual(values = brewer.pal(8, "Accent")) +
  guides(fill = guide_colorbar(title.position = "top")) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12)
  )
```
Didn't really work 


# Metabolite network

## Subset variable features
```{r}
Area_long_batch_highCV <- Area_long_batch %>% 
  filter(feature %in% Common_features$feature) %>% 
  group_by(feature) %>% 
  summarize(CV = sd(log_Area)/mean(log_Area)) %>% 
  mutate(cutoff = quantile(CV, 0.75)) %>%
  filter(CV > cutoff) %>% 
  ungroup()

head(Area_long_batch_highCV)
dim(Area_long_batch_highCV)
```

```{r}
feature_by_cell_highCV <- Area_long_batch %>% 
  filter(feature %in% Area_long_batch_highCV$feature) %>% 
  select(cell, feature, z) %>% 
  spread(feature, z) %>% 
  as.data.frame()

row.names(feature_by_cell_highCV) <- feature_by_cell_highCV$cell

head(feature_by_cell_highCV)

cor_met_mat <- cor(feature_by_cell_highCV[,-1],
                   method = "s")
dim(cor_met_mat)
```

## Make edge table 
```{r}
cor_met_mat2 <- cor_met_mat
cor_met_mat2[lower.tri(cor_met_mat2, diag = T)] <- NA

number_of_cells <- ncol(Area_z_wide_batch) - 1
```

```{r}
edge_table <- cor_met_mat2 %>% 
  as.data.frame() %>% 
  mutate(from = row.names(cor_met_mat2)) %>% 
  gather("to", "r", 1:ncol(cor_met_mat2)) %>% 
  filter(is.na(r) == F) %>% 
  mutate(t = r * sqrt((number_of_cells-2)/(1-r^2))) %>% 
  mutate(p.value = case_when(
    t > 0 ~ pt(t, df = number_of_cells-2, lower.tail = F),
    t <=0 ~ pt(t, df = number_of_cells-2, lower.tail = T)
  )) %>% 
  mutate(FDR = p.adjust(p.value, method = "fdr")) %>% 
  mutate(Type = case_when(
    r > 0.8 ~ "sig",
    T ~ "not"
  )) %>% 
  filter(r > 0) %>% 
  filter(Type == "sig")


head(edge_table)
```
### Are known compounds there? 
```{r}
edge_table %>% 
  filter(r > 0.8) %>% 
  arrange(-abs(r)) %>% 
  filter(str_detect(from, "[a-z]") |
           str_detect(to, "[a-z]"))
```
## Make node table

* Compound name, 
* Most correlated PC 

```{r}
Node_table <- feature_by_PC_df %>% 
  filter(index <= 25) %>% 
  group_by(feature) %>% 
  slice_max(n = 1, order_by = abs(r)) %>% 
  arrange(index) %>% 
  rename(max.index = index) %>% 
  rename(represented.by = PC) %>% 
  mutate(Cell = case_when(
    represented.by == "PC5" ~ "idioblast",
    T ~ "other"
  )) %>% 
  ungroup() %>% 
  filter(feature %in% edge_table$from |
           feature %in% edge_table$to) %>% 
  select(feature, represented.by, max.index, Cell) %>% 
  mutate(represented.by = reorder(represented.by, max.index)) %>% 
  mutate(tag2 = case_when(
    str_detect(feature, "[a-z]") ~ "known",
    T ~ ""
  )) %>% 
  mutate(tag3 = case_when(
    str_detect(feature, "[a-z]") ~ feature
  ))

head(Node_table)
```
## Graph network 
```{r}
metabolite_network <- graph_from_data_frame(edge_table, vertices = Node_table, directed = F)

metabolite_network_layout <- layout_with_kk(metabolite_network)
```

```{r}
ggraph(metabolite_network, layout = "kk", circular = F) +
  geom_edge_diagonal(color = "grey70", width = 0.5, alpha = 0.5) +
  geom_node_point(alpha = 0.8, color = "white", shape = 21,
                  aes(size = tag2, fill = Cell)) + 
  geom_node_text(aes(label = tag3), size = 3, repel = T, max.overlaps = 10) +
  scale_edge_color_manual(values = c("steelblue3", "tomato1")) +
  scale_fill_manual(values = c("slateblue3", "grey80"),
                    limits = c("idioblast", "other")) +
  scale_size_manual(values = c(1.2, 3),
                    limits = c("", "known")) +
  labs(fill = NULL) +
  guides(size = "none",
         fill = guide_legend(override.aes = list(size = 4), 
                             title.position = "top", nrow = 1)) +
  theme_void()+
  theme(
    text = element_text(size = 14), 
    legend.position = "bottom",
    legend.justification = 1,
    title = element_text(size = 12)
  )
```
 

